---
author: "John Little"
---

# Get ACS Census data

## Load library packages

```{r}
#| warning: false
#| message: false
#| eval: false
library(tidyverse)
library(tidycensus)
```

### More information about packages

-   [{tidycensus}](https://walker-data.com/tidycensus/index.html)
-   [*Analyzing US Census Data: Methods, Maps, and Models in R* by Kyle Walker](https://walker-data.com/census-r)
-   [{tidyverse}](https://tidyverse.org)

## Get variable names for 2020 ACS census

```{r}
#| eval: false
my_vars <- tidycensus::load_variables(2020, "acs5")
```

## Filter code book

The ACS data element is Total population. `B01003_001`

```{r}
#| eval: false
my_vars |> 
  filter(name == "B01003_001")
```

## Get data

The data were downloaded via the {tidycensus} package, using the `get_acs()` function. Limiting the data to ACS data for the year 2020, further limited to *county* census geography for the state of *North Carolina*. By setting the argument `geometry` to *TRUE*, we download Census Tigerline data in the shapefile format.

### More information

ACS - American Community Surveys - [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/)

TIGER/Line shapefiles - [TIGER/Line Shapefiles](https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html) - [TIGER/Line Files Technical Documentation](https://www.census.gov/programs-surveys/geography/technical-documentation/complete-technical-documentation/tiger-geo-line.html)

```{r}
#| eval: false
my_data_and_shapes <- 
  tidycensus::get_acs("county", 
                      variables = "B01003_001",
                      year = 2020,
                      state = "NC", 
                      geometry = TRUE)
```

## Preview data

```{r}
#| eval: false
my_data_and_shapes |> 
  as_tibble() 
```

## Save data

```{r}
#| eval: false
sf::st_write(my_data_and_shapes, "data/raw/ACS_nc-county_shapefile.shp")
```

## Data Dictionary

-   *variable* - ACS census variable

-   *estimate* - ACS census variable estimate

-   *moe* - ACS census variable margin of error

-   *geometry* - Census TIGER/Line shapefile
